---
title: "Mini-Pitch for ShinyApp"
author: "git-comp"
date: "24 5 2021"
output: ioslides_presentation
widescreen: yes
smaller: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Context, Data, Sources
### Context
For this assignment, I have used **official COVID-19 data for Switzerland** available as of May 21, 2021. I have generated an interactive overview of COVID-19 cases per calender week from Feb 2020 until May 2021 per State (called Canton in Switzerland).

In the ShinyUI, checkboxes allow the selection of individual states. In the pitch presentation, I have selected all states.

### Data Source
https://www.covid19.admin.ch/api/data/20210521-abejzev6/downloads/sources-csv.zip

### Overview of Structure and further info
https://www.covid19.admin.ch/en/overview

## Data Preparation (I/II)
```{r echo = TRUE, cache= TRUE}
library("data.table")
library("ggplot2")
fileUrl <- "https://www.covid19.admin.ch/api/data/20210521-abejzev6/downloads/sources-csv.zip" 
download.file(fileUrl, destfile = paste0(getwd(), '/sources-csv.zip'), method = "curl")
unzip("sources-csv.zip")

## parse file with weekly figures per georegion (incl. aggregated 
##  regions CH01..CH07, CHFL, CH)
casesGeoRegionDT <- data.table::fread(input = ".//data//COVID19Cases_geoRegion_w.csv")

## Only select columns geoRegion, datum, entries
cpgraw <- casesGeoRegionDT[, c("geoRegion","datum","entries")]
```

## Data Preparation (II/II)
```{r echo = TRUE, cache= TRUE}
## remove aggregated regions CH01..CH07, CHFL, CH to avoid redundancy and double counting
cpgraw <- cpgraw[cpgraw$geoRegion != "CH" & cpgraw$geoRegion != "CH01" & 
                   cpgraw$geoRegion != "CH02" & cpgraw$geoRegion != "CH03" & 
                   cpgraw$geoRegion != "CH04" & cpgraw$geoRegion != "CH05" & 
                   cpgraw$geoRegion != "CH06" & cpgraw$geoRegion != "CH07" & 
                   cpgraw$geoRegion != "CHFL"]

cpgraw$geoRegionFactor <- as.factor(cpgraw$geoRegion)
cpgraw$datumFactor <- as.factor(cpgraw$datum)

cpgrawDF <- as.data.frame(cpgraw)

```

## COVID-19 cases in Switzerland per region and calendar week
```{r echo = FALSE, cache= TRUE, warning=FALSE}
## In the app, filters would now be applied based on checkboxes, e.g. 
# cpgrawDFDisplay <- rbind(cpgrawDFDisplay, subset(cpgrawDF,cpgrawDF$geoRegion == "ZG"))

## For the presentation, we add all States/Cantons
cpgrawDFDisplay <- cpgrawDF

ggplot(cpgrawDFDisplay, aes(x = datumFactor, y=entries)) +
            geom_col(aes(fill = geoRegionFactor))

```

